---
title: "Gapminder HW"
author: "Bilal Gilani"
date: "3/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1

## 1.
```{r}
if(!file.exists("./data")) {dir.create("./data")}

fileURLs <- c("https://docs.google.com/spreadsheet/pub?key=0AkBd6lyS3EmpdHo5S0J6ekhVOF9QaVhod05QSGV4T3c&output=xlsx", 
              "https://docs.google.com/spreadsheet/pub?key=phAwcNAVuyj2tPLxKvvnNPA&output=xlsx", 
              "https://docs.google.com/spreadsheet/pub?key=tSUr_yZVbM6a3AGJEq_Z2Pw&output=xlsx",
              "https://docs.google.com/spreadsheet/pub?key=0ArfEDsV3bBwCdHBzUVVSMDlTX1ZCUnNJQ3ZFdkFXVFE&output=xlsx",
              "https://docs.google.com/spreadsheet/pub?key=phAwcNAVuyj0XOoBL_n5tAQ&output=xlsx" ) 

var_names <- c("GDP","life_expectancy", "alt_GDP", "blood press", "population")
```



```{r}
library(readxl)
library(tidyr)
library(tidyverse)
get_clean <- function(url_in, var_name) {
  download.file(url_in, destfile = "./data/tmp.xslx", mode = "wb")
  output <- read_excel("./data/tmp.xslx")
  names(output)[[1]] <- "country"
  output <- output %>%
    pivot_longer(-country, names_to = "year", values_to = var_name) %>%
    mutate(year = as.numeric(year)) %>%
    drop_na()
}
out1 <- get_clean(fileURLs[1], var_names[1])
head(out1)

```

```{r}
all_data <- map2(fileURLs, var_names, get_clean)

```

## 2.
```{r}
joined_data <- all_data %>%
  reduce(full_join, by = c("country", "year"))
```

## 3.
```{r}
library(countrycode)
continent <- countrycode(sourcevar = joined_data$country, origin = "country.name", destination = "continent")
new_gapminder <- cbind(joined_data, continent)

no_continent <- new_gapminder[is.na(new_gapminder$continent),]
unique(no_continent$country)

## These countries were not given values in the continent column because they no longer exist
## or are islands such as in the examples of the Channel Islands and St. Martin, or territories of 
## another country such as "Akrotiri and Dhekelia" which is a British Overseas Territory and considered
## part of the continent however is not part of the continent shelf or Cocos Island which is an Australian
## territory but is considered to be a part of Asia. Another possibility is that they were considered duplicates
## such a in the case of "Eritrea and Ethiopia", both of which are now individual countries.

new_gapminder$continent[which(new_gapminder$country == "Channel Islands")] <- "Europe"
new_gapminder$continent[which(new_gapminder$country == "Kosovo")] <- "Europe"
new_gapminder$continent[which(new_gapminder$country == "Serbia and Montenegro")] <- "Europe"
new_gapminder$continent[which(new_gapminder$country == "Central African Rep.")] <- "Africa"
new_gapminder$continent[which(new_gapminder$country == "Akrotiri and Dhekelia")] <- "Europe"
new_gapminder$continent[which(new_gapminder$country == "Cocos Island")] <- "Asia"
new_gapminder$continent[which(new_gapminder$country == "Czechoslovakia")] <- "Europe"
new_gapminder$continent[which(new_gapminder$country == "East Germany")] <- "Europe"
new_gapminder$continent[which(new_gapminder$country == "Eritrea and Ethiopia")] <- "Africa"
new_gapminder$continent[which(new_gapminder$country == "St. Martin")] <- "Americas"
new_gapminder$continent[which(new_gapminder$country == "North Yemen (former)")] <- "Asia"
new_gapminder$continent[which(new_gapminder$country == "South Yemen (former)")] <- "Asia"
new_gapminder$continent[which(new_gapminder$country == "Yugoslavia")] <- "Europe"

which(is.na(new_gapminder$continent))

## no missing values in the continent column.

new_gapminder <- new_gapminder %>%
  arrange(country, year)

```

## 4.
```{r}
a <- new_gapminder[,c(1,2,3,4,7,8)]
a <- a %>% fill(population)
plot_data <- a %>% distinct()



```


```{r}
library(ggplot2)
library(gganimate)
library(gifski)

plot1 <- ggplot(plot_data, aes(GDP, life_expectancy)) +
  geom_point(alpha = 0.7, aes(size = population, colour = continent)) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  theme(legend.position = "right") +
  labs(title = "Year: {frame_time}", x = "GDP", y = "life expectancy") +
  transition_time(year) +
  ease_aes("linear")

animate(plot1)

```

# Part 2

## 1.
```{r}
attach(plot_data)
plot(GDP, life_expectancy) +
  abline(lm(life_expectancy ~ GDP))
```

## 2.
```{r}
reg <- lm(life_expectancy ~ GDP, data = plot_data)
res.lm <- resid(reg)
NoNA <- na.omit(plot_data)
NoNA <- cbind(NoNA, res.lm)

plot2 <- ggplot(NoNA, aes(GDP, res.lm)) +
  geom_point() +
  xlab("GDP") +
  ylab("Residuals") +
  facet_grid(rows = vars(continent))
  
plot2
```

## 3.
```{r}
library(broom)
glance(reg)
plot(reg)

## The outliers are in rows 35,443 , 45,009 , and 45,012.

outlier.reg <- plot_data %>%
  slice(35443, 45009, 45012)

head(outlier.reg)
```
The outliers in this model are Rwanda in 1994 which can be attributed to the Rwandan genocide. The other
two outliers is the UAE in both 1977 and 1980. I could not find significant events as to why the UAE stuck
out in the particular years.

# Part 3

```{r}
library(gapminder)
data("gapminder")
```

## a.
```{r}
gapminder %>%
  mutate(year = year - mean(year)) ->
  gapminder

mean(gapminder$year)

```


## b.
```{r}
attach(gapminder)
year2 <- (year)^2
gapminder <- cbind(gapminder, year2)

quad <- lm(lifeExp ~ year + year2, data = gapminder)
quad
summary(quad)

```
The quadratic formula is Y = 60.512 + 0.33X - 0.003X^2 with X being the year and Y being life expectancy.


## c.
```{r}
glance(quad)
plot(quad)

## The outliers are in rows 222 , 1,293 , and 1,294.

outlier.quad <- gapminder %>%
  slice(222, 1293, 1294)

head(outlier.quad)

```
Once again, Rwanda is an outlier, which can be attributed to the genocide. Cambodia in the year 1977 is
an outlier in this model. Upon further research I found that there was a genocide in Cambodia in 1977, which
could be the reason for its outlier status.











